---
title: "bigKahuna - all the maps in the challenge"
author: "Timothy Oliver"
date: "2023-10-31"
output:
  html_document:
    toc: true
    toc_float: true
    code_folding: hide
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r please}
library(tidycensus)
library(tidyverse)
library(scales)
library(tmap)
library(sf)
library(mapsf)
library(socviz)
## gg-suite
library(ggplot2)
library(ggbeeswarm)
library(ggiraph)
library(ggtext)
library(ggspatial)
library(ggmap)

## google api key for ggmap
# ggmap::register_google(key=)

source("D:/timol/Grad_School/plotTheme_TO.R")

options(scipen = 999)
```


```{r get_areas,results='hide'}

us_area <- get_decennial(
  geography = "state",
  year = 2010,
  variables = "P001001",
  geometry = T) %>% st_transform("EPSG:3665") 

cont_us_area <- us_area %>%
  filter(NAME != "Alaska" & NAME != "Hawaii" & NAME != "Puerto Rico")


```


# 1	Points	A map with points

```{r data_prep, results='hide'}
## earthquakes to hurricanes?
# from usgs earthquake database
earthquakes_23 <- st_read("data/2023_US_earthquakes.json") %>% st_transform(st_crs(us_area))

inBoundsEQ <- st_join(earthquakes_23, cont_us_area, join=st_within)[cont_us_area,]
# ensure only points in continental us

# View(inBoundsEQ %>% group_by(NAME))
test <- inBoundsEQ %>% group_by(NAME) %>% summarize(stateMaxMag = max(mag),
                                                    stateMinMag = min(mag))

inBoundsEQ <- st_join(x=inBoundsEQ,y=test)
inBoundsEQ <- inBoundsEQ %>% mutate(stateMax =ifelse(mag==stateMaxMag,1,0),
                                 stateMin = ifelse(mag==stateMinMag,1,0),
                                 tooltip = paste(paste("Maximum",NAME.x,"Magnitude"), mag, sep = ": "),
                                 color = ifelse(mag==max(inBoundsEQ$mag),RED2,RED1))
# magMax <- max(inBoundsEQ$mag)
```


```{r points}

pointMap <- ggplot()+
  # annotation_map_tile(type="loviniacycle",data=cont_us_area)+ # adds annotation, but lose projection without data
  # ggmap(get_map(maptype = 'terrain-background', source = 'stamen')) # needs api key
  geom_sf(data = cont_us_area, fill = "lightgrey",col = "white")+
  geom_sf(data = inBoundsEQ[inBoundsEQ$stateMin==0 & inBoundsEQ$stateMax==0,],
          color=GRAY7)+
  # geom_sf(data = inBoundsEQ[inBoundsEQ$stateMin==1,],
  #         color="yellow")+
  geom_sf_interactive(data = inBoundsEQ[inBoundsEQ$stateMax==1,],
                      color=RED1,
                      aes(tooltip = tooltip))+
  geom_sf_interactive(data = inBoundsEQ[inBoundsEQ$mag==max(inBoundsEQ$mag),],
                      color="#FA5128",
                      aes(tooltip = tooltip))+
  # geom_sf_text(data = inBoundsEQ[inBoundsEQ$mag==max(inBoundsEQ$mag),], # or just the absolute max: inBoundsEQ$mag==max(inBoundsEQ$mag); originally all maxes: inBoundsEQ[inBoundsEQ$stateMax==1,]
  #              aes(label = mag, face = "bold"),
  #              color = RED1, vjust = -1.5, hjust = -1)+
  labs(caption = "Source | USGS Earthquakes Catalog\nTimothy Oliver",
       title = "2023 - Year of Minor Intensity Earthquakes",
       subtitle= "<span style='font-size:11pt'>Highest Recorded at
       <span style='color:#FA5128;'>5.48</span> in 
       <span style='color:#FA5128;'>California</span>")+
  theme_test()+
  theme(axis.title.x = element_blank(),
        axis.text.x = element_blank(),
        axis.title.y = element_blank(),
        axis.text.y = element_blank(),
        # axis.ticks.x = element_blank(),
        # axis.ticks.y = element_blank(),
        panel.background = element_rect(
        fill = "#fcfcf4",
        colour = NA),
        # allows html in labs
        plot.subtitle = element_markdown(lineheight = 1.1,hjust = 0)
        )

girafe(ggobj = pointMap) %>%
  girafe_options(opts_hover(css = "fill:cyan;"), 
                 opts_zoom(max = 10))

## idea: highest and lowest magnitude earthquakes by state

ggsave(filename="maps/StatesHighestMagnitudeEarthquakes_2023.png",units="cm",width=24,height=18,dpi=150,device="png")
```

# 2	Lines	A map with lines

```{r lines}

```


# 3	Polygons	A map with polygons

# 4	A bad map	Let's get this over with

# 5	Analog Map	Non-digital maps

# 6	Asia	Largest of the continents

# 7	Navigation	A map that helps you to navigate

# 8	Africa	Second-largest and second-most populous continent

# 9	Hexagons	6 sides, 6 angles, and 6 vertices

# 10	North America	Northern part of the American continent

# 11	Retro	A blast from the past

# 12	South America	Southern part of the American continent

# 13	Choropleth	Classic thematic map: a choropleth

# 14	Europe	The westernmost peninsulas of Eurasia

# 15	OpenStreetMap	The greatest of the datasets. Remember to give credit.

# 16	Oceania	Region made up of thousands of islands throughout the Central and South Pacific

# 17	Flow	Flow of transport, people

# 18	Atmosphere	Conditions of the atmosphere can be either weather or climate

# 19	5-minute map	Spend no more than 5 minutes to prepare this map

# 20	Outdoors	Map of mountains, trails, or something completely different

# 21	Raster	Pixels, please

# 22	North is not always up	⬆️

# 23	3D	The third dimension

# 24	Black & white	Only two colors allowed

# 25	Antarctica	A cold continent that can't be reached with Web Mercator

# 26	Minimal	Less is more

# 27	Dot	Dot density, a single dot in space or something different

# 28	Is this a chart or a map?	In thematic maps, you can't always tell. Try to stretch the limits

# 29	Population	A classic theme for a map

# 30	"My favorite.."	You choose!

```{r knit_etc., eval=F, include=F}
rmarkdown::render(input="all_maps.Rmd",output_format = "html_document",output_file = "The Thirty-Day Mapping Journey.html")
```