---
title: "bigKahuna - all the maps in the challenge"
author: "Timothy Oliver"
date: "2023-10-31"
output:
  html_document:
    toc: true
    toc_float: true
    code_folding: hide
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE,results='hide',message=F,warning=F)
```

```{r please}
library(tidycensus)
library(tidyverse)
library(scales)
library(tmap)
library(sf)
library(sfheaders)
library(mapsf)
library(socviz)
library(sp)
library(gstat)
## gg-suite
library(ggplot2)
library(ggbeeswarm)
library(ggiraph)
library(ggtext)
library(ggspatial)
library(ggmap)
library(jsonlite)
library(httr)
library(osmdata)

## google api key for ggmap
# ggmap::register_google(key=)

source("themes/plotTheme_TO.R")

options(scipen = 999)
```


```{r get_areas,results='hide'}

us_area <- get_decennial(
  geography = "state",
  year = 2010,
  variables = "P001001",
  geometry = T) %>% st_transform("EPSG:3665") 

cont_us_area <- us_area %>%
  filter(NAME != "Alaska" & NAME != "Hawaii" & NAME != "Puerto Rico")

PA <- cont_us_area[cont_us_area$NAME=="Pennsylvania",]

philly <- get_decennial(
  geography = "tract",
  state="PA",
  county="Philadelphia",
  year = 2010,
  variables = "P001001",
  geometry = T,
  output="wide") %>% st_transform("EPSG:2272") 




```

```{r data_test, results='hide'}
birds <- st_read("data/birds_gps/gps/anon_gps_tracks_with_dive.csv",options=c("X_POSSIBLE_NAMES=x","Y_POSSIBLE_NAMES=y")) %>%
  st_as_sf(., coords = c("lon", "lat"), crs = 4326)

obesity <- st_read("data/National_Obesity_By_State-shp/National_Obesity_By_State.shp") %>% st_transform(st_crs(us_area))

ocean <- st_read("data/ocean/ne_110m_ocean.shp") %>% st_transform(crs=4326)

chile <- st_read("data/chl_adm_bcn_20211008_shp/chl_admbnda_adm0_bcn_20211008.shp")

russia <- st_read("data/russia/bs331gg4381.shp")

canada <- st_read("data/canada/gpr_000b11a_e.shp")

waterway <- st_read("data/Navigable_Waterway_Network_Lines/Waterway_network.shp")

dead_sea <- st_read("data/deadsea.csv",options=c("X_POSSIBLE_NAMES=x","Y_POSSIBLE_NAMES=y")) %>%
  mutate(x = LON,y=LAT) %>%
  st_as_sf(., coords = c("LON", "LAT"), crs = 4326) %>% mutate(elev = 0-as.numeric(DEPTH))

deadSea <- read_csv("data/deadsea.csv") %>%
  st_as_sf(., crs = "EPSG:61416405", coords = 
    c("LON", "LAT")) 
# %>% st_set_crs("EPSG:61416405") |>    ?
    # st_transform(4141) 

paParks <- st_read("data/Pennsylvania_State_Park_Boundaries/Pennsylvania_State_Park_Boundaries.shp")


# global_water <- st_read("data/globalWater/goas_v01.shp") # not working
```

```{r other_data}
# # wlaRoads <- st_read('data/w_LA_roads.geojson')
# # laTrails <- st_read('data/LA_pedestrianLines.geojson')
# osm4 <- opq(bbox = "caddo parish") %>%
#   add_osm_feature(key = "highway") %>%
#   osmdata_sf()
# 
# 
# la <- 
#   get_acs(geography = "county", 
#           variables = c("B25026_001E",
#                         "B02001_002E",
#                         "B19013_001E",
#                         "B25058_001E",
#                         "B06012_002E"
#                         ), 
#           year=2020, state="LA", 
#           geometry=TRUE, output="wide") %>%
#   st_transform('EPSG:2027') %>%
#   rename(TotalPop = B25026_001E,
#          Whites = B02001_002E,
#          MedHHInc = B19013_001E,
#          MedRent = B25058_001E,
#          TotalPoverty = B06012_002E
#          ) %>%
#   dplyr::select(-NAME) %>%
#   mutate(pctWhites = ifelse(TotalPop > 0, Whites / TotalPop,0),
#          pctPoverty = ifelse(TotalPop > 0, TotalPoverty / TotalPop, 0),
#          # pctVacant = ifelse(TotalHHs > 0, VacantHHs / TotalHHs, 0),
#          # avgHHsizes = ifelse(TotalHHs > 0, TotalPop / TotalHHs, 0),
#          year = "2020") %>%
#   dplyr::select(-Whites, -TotalPoverty, -(ends_with("E")),-(ends_with("M"))) 
# 
# # trail_to_users <- st_join(laTrails,la,intersection=st_within)
# # 
# # trail_to_users <- trail_to_users %>% group_by(GEOID) %>% summarize(trail_len = sum(length)) %>% mutate(walker_ratio = TotalPop/trail_len)
# 
# road_to_users <- st_join(osm4$osm_lines %>% st_transform(st_crs(la)),la,join=st_within)
# 
# road_to_users <- road_to_users %>% group_by(GEOID,TotalPop) %>% tally() %>% mutate(driver_ratio = TotalPop/n)
```

# 1	Points	A map with points

```{r data_prep, results='hide'}
## earthquakes to hurricanes?
# from usgs earthquake database
earthquakes_23 <- st_read("data/2023_US_earthquakes.json") %>% st_transform(st_crs(us_area))

inBoundsEQ <- st_join(earthquakes_23, cont_us_area, join=st_within)[cont_us_area,]
# stateEQ <- st_join(cont_us_area, earthquakes_23,join=st_contains)#[earthquakes_23,]
stateEQ <- st_join(cont_us_area,st_as_sf(over(as_Spatial(cont_us_area),as_Spatial(earthquakes_23))),join=st_contains)
# ensure only points in continental us

# View(inBoundsEQ %>% group_by(NAME))
test <- inBoundsEQ %>% group_by(NAME) %>% summarize(stateMaxMag = max(mag),
                                                    stateMinMag = min(mag))
# test2 <- stateEQ %>% group_by(NAME) %>% tally %>% rename(eqCount = n)

inBoundsEQ <- st_join(x=inBoundsEQ,y=test)
# stateEQ <- st_join(x=stateEQ,y=test2)
inBoundsEQ <- inBoundsEQ %>% mutate(stateMax =ifelse(mag==stateMaxMag,1,0),
                                 stateMin = ifelse(mag==stateMinMag,1,0),
                                 tooltip = paste(paste("Maximum",NAME.x,"Magnitude"), mag, sep = ": "),
                                 color = ifelse(mag==max(inBoundsEQ$mag),RED2,RED1))

states= st_join(cont_us_area,inBoundsEQ %>% group_by(NAME.x) %>% tally())
# magMax <- max(inBoundsEQ$mag)
```


```{r points}

pointMap <- ggplot()+
  # annotation_map_tile(type="loviniacycle",data=cont_us_area)+ # adds annotation, but lose projection without data
  # ggmap(get_map(maptype = 'terrain-background', source = 'stamen')) # needs api key
  geom_sf(data = cont_us_area, fill = "lightgrey",col = "white")+
  geom_sf(data = inBoundsEQ[inBoundsEQ$stateMin==0 & inBoundsEQ$stateMax==0,],
          color=GRAY7)+
  # geom_sf(data = inBoundsEQ[inBoundsEQ$stateMin==1,],
  #         color="yellow")+
  geom_sf_interactive(data = inBoundsEQ[inBoundsEQ$stateMax==1,],
                      color=RED1,
                      aes(tooltip = tooltip))+
  geom_sf_interactive(data = inBoundsEQ[inBoundsEQ$mag==max(inBoundsEQ$mag),],
                      color="#FA5128",
                      aes(tooltip = tooltip))+
  # geom_sf_text(data = inBoundsEQ[inBoundsEQ$mag==max(inBoundsEQ$mag),], # or just the absolute max: inBoundsEQ$mag==max(inBoundsEQ$mag); originally all maxes: inBoundsEQ[inBoundsEQ$stateMax==1,]
  #              aes(label = mag, face = "bold"),
  #              color = RED1, vjust = -1.5, hjust = -1)+
  labs(caption = "Source | USGS Earthquakes Catalog\nTimothy Oliver",
       title = "2023 - Year of Minor Intensity Earthquakes",
       subtitle= "<span style='font-size:11pt'>Highest Recorded at
       <span style='color:#FA5128;'>5.48</span> in 
       <span style='color:#FA5128;'>California</span>")+
  theme_test()+
  theme(axis.title.x = element_blank(),
        axis.text.x = element_blank(),
        axis.title.y = element_blank(),
        axis.text.y = element_blank(),
        # axis.ticks.x = element_blank(),
        # axis.ticks.y = element_blank(),
        panel.background = element_rect(
        fill = "#fcfcf4",
        colour = NA),
        # allows html in labs
        plot.subtitle = element_markdown(lineheight = 1.1,hjust = 0)
        )

girafe(ggobj = pointMap) %>%
  girafe_options(opts_hover(css = "fill:cyan;"), 
                 opts_zoom(max = 10))

## idea: highest and lowest magnitude earthquakes by state

ggsave(filename="maps/StatesHighestMagnitudeEarthquakes_2023.png",units="cm",width=24,height=18,dpi=150,device="png")
```

# 2	Lines	A map with lines

```{r lines}

q <- opq(bbox = "parque o'higgins, santiago") %>%
    add_osm_feature(key = 'highway') %>% 
  osmdata_sf()

ggplot()+
  geom_sf(data=q$osm_polygons)+
  geom_sf(data=q$osm_lines)+
  # scale_y_reverse()+
  theme_minimal()+
  theme(panel.grid = element_blank(),
        axis.text= element_blank(),
        panel.background = element_rect(
        fill = "#fcfcf4",
        colour = NA))+
  ggspatial::annotation_north_arrow(
    location = "tr", which_north = "true",
    pad_x = unit(0.75, "in"), pad_y = unit(3, "in"),
    style = ggspatial::north_arrow_nautical(
      line_col = "grey20",
      text_family = "ArcherPro Book"
    )
  )+
  labs(title="74 Hectares of Green",subtitle="Parque o'Higgins - Santiago, Chile",
       caption = "Source | OpenStreetMaps - Overpass API")

ggsave(filename="maps/02-lines.png",units="cm",width=24,height=18,dpi=150,device="png")

```

```{python trial, include = F, eval = F}
# has new environment data
print(f"trying this out\n{RED1} vs {RED2}")
```


# 3	Polygons	A map with polygons


# 4	A bad map	Let's get this over with
```{r bad_Alaska}
orig <- cbind(150.167,65.833) %>% cbind( ., 1:nrow(.) )
dest <- cbind(180.167,65.833) %>% cbind( ., 1:nrow(.) )
# order by id in case you have more than two points
m <- rbind(orig,dest) %>% .[order(.[,3]),]

# make lines
sfc <- sfheaders::sfc_linestring(
  obj = m, linestring_id = 3)


# center of Bering strait Land Bridge
# 65°50′N 164°10′W

ggplot()+
  geom_sf(data=us_area%>%filter(NAME!="Puerto Rico"&NAME!="Hawaii")%>%st_transform(crs=2964))+
  # geom_sf(data=canada%>%st_transform(crs=2964))+
  geom_sf(data=sfc%>%st_set_crs(4326)%>%st_transform(crs=2964),color=GRAY6)+
  theme(panel.grid = element_blank(),
        axis.text= element_blank(),
        panel.background = element_rect(
        fill = "#fcfcf4",
        colour = NA))+
  labs(title="Route to the Americas - The Bering Strait",subtitle="A Beginning")
  
  
ggsave(filename="maps/04-badMap.png",units="cm",width=24,height=18,dpi=150,device="png")
```

# 5	Analog Map	Non-digital maps

# 6	Asia	Largest of the continents

# 7	Navigation	A map that helps you to navigate

```{r boardwalk_data}
boardwalks <- st_read("data/Boardwalk/geo_export_b0a28836-8305-48bf-aae6-54f6ed86af65.shp") %>% st_transform(st_crs(cont_us_area))

ggplot()+
  geom_sf(data = cont_us_area[cont_us_area$NAME == "New Jersey",])+
  geom_sf(data = boardwalks, color = "blue", linewidth=0.75)
```

```{r navigation}
navMap <- ggplot()+
  geom_sf(data = PA, fill = "lightgrey",col = "white")+
  geom_sf(data = inBoundsEQ[inBoundsEQ$stateMin==0 & inBoundsEQ$stateMax==0,],
          color=GRAY7)+
  # geom_sf(data = inBoundsEQ[inBoundsEQ$stateMin==1,],
  #         color="yellow")+
  geom_sf_interactive(data = inBoundsEQ[inBoundsEQ$stateMax==1,],
                      color=RED1,
                      aes(tooltip = tooltip))+
  geom_sf_interactive(data = inBoundsEQ[inBoundsEQ$mag==max(inBoundsEQ$mag),],
                      color="#FA5128",
                      aes(tooltip = tooltip))+
  # geom_sf_text(data = inBoundsEQ[inBoundsEQ$mag==max(inBoundsEQ$mag),], # or just the absolute max: inBoundsEQ$mag==max(inBoundsEQ$mag); originally all maxes: inBoundsEQ[inBoundsEQ$stateMax==1,]
  #              aes(label = mag, face = "bold"),
  #              color = RED1, vjust = -1.5, hjust = -1)+
  labs(caption = "Note: earthquakes of magnitude 2.5 or greater shown\n\nSource | USGS Earthquakes Catalog\nTimothy Oliver",
       title = "2023 - A Year of Minor Intensity Earthquakes",
       subtitle= "<span style='font-size:11pt'>Highest Recorded at
       <span style='color:#FA5128;'>5.48</span> in 
       <span style='color:#FA5128;'>California</span>")+
  theme_test()+
  theme(axis.title.x = element_blank(),
        axis.text.x = element_blank(),
        axis.title.y = element_blank(),
        axis.text.y = element_blank(),
        # axis.ticks.x = element_blank(),
        # axis.ticks.y = element_blank(),
        panel.background = element_rect(
        fill = "#fcfcf4",
        colour = NA),
        # allows html in labs
        plot.subtitle = element_markdown(lineheight = 1.1,hjust = 0)
        )

ggsave(filename="maps/7-navigation.png",units="cm",width=24,height=18,dpi=150,device="png")
```


# 8	Africa	Second-largest and second-most populous continent

# 9	Hexagons	6 sides, 6 angles, and 6 vertices

Checkout [Andrew Wheeler's](https://andrewpwheeler.com/2019/08/07/making-a-hexbin-map-in-ggplot/) quick write-up about making a hex-bin map completely in R.

```{r hex} 
# Functions and such from Andrew Wheeler
height = 750
lev_cnt <- seq(0,225,25)

wd_hex <- function(height){
  tri_side <- height/2
  sma_side <- height/4
  width <- 2*sqrt(tri_side^2 - sma_side^2)
  return(width)
}

hex_dim <- function(area){
  num <- 4*area
  den <- 6*sqrt(3)
  vert <- 2*sqrt(num/den)
  horz <- wd_hex(height)
  return(c(vert,horz))
}

my_dims <- hex_dim(1000^2)   #making it a square kilometer

ggplot()+
  geom_hex(data=inBoundsEQ, color='white', alpha=0.85, size=0.1, binwidth=my_dims, 
                        aes(x=830000,y=72300,fill=cut(..count..,lev_cnt))) + 
               scale_fill_brewer(name="Count Earthquakes", palette="OrRd")
```

# 10	North America	Northern part of the American continent

# 11	Retro	A blast from the past

# 12	South America	Southern part of the American continent

# 13	Choropleth	Classic thematic map: a choropleth

```{r gitAPI, include=T,echo=T,results='hide'}
us_area <- get_decennial(
  geography = "state",
  year = 2010,
  variables = "P001001",
  geometry = T) %>% st_transform("EPSG:3665") %>%
  filter(NAME != "Puerto Rico")

# cont_us_area <- us_area %>%
#   filter(NAME != "Alaska" & NAME != "Hawaii" & NAME != "Puerto Rico")

stateUsers = c()

## Get the state names and place results of the following in a dataframe for reference
### divide total count by population and make grad-symbol map or similar plot/map 
#### Puerto Rico has space, so we won't use it (also territory)

# uses month to cut day, change if not November
currentTime = paste("Nov",strsplit(date(),"Nov ")[[1]][2])

library(httr)
for (state in us_area$NAME){
    # prep
  state <- tolower(gsub(" ","",state))
  count <- fromJSON(rawToChar((GET(paste("https://api.github.com/search/users?q=location:",state,sep=""))$content)))$total_count
  
   # sleep to minimize chance of NULL
  if (length(stateUsers)%%8==0){
  Sys.sleep(10)}
  
   # get counts
  if (is.null(count)){
    stateUsers <- c(stateUsers,NA)
  }else{
  stateUsers <- c(stateUsers,count)
  }
}

us_area$userCount <- stateUsers
rm(stateUsers,count)
us_area$userStandard <- us_area$userCount/us_area$value*100

```

```{r choro_map}
ggplot(data = us_area %>% filter(NAME != "Alaska" & NAME != "Hawaii"))+
  geom_sf(aes(fill=userStandard),color="white")+
  scale_fill_binned()+
  labs(title = paste("Active Github Users as of\n",currentTime))+
  guides(fill=guide_legend(title="Percent of Population\nActively Using\nGithub"))+
  theme_test()+theme(axis.text.x=element_blank(),axis.text.y=element_blank())

ggsave(filename="maps/13-choro.png",units="cm",width=24,height=18,dpi=150,device="png")
  
```


# 14	Europe	The westernmost peninsulas of Eurasia

# 15	OpenStreetMap	The greatest of the datasets. Remember to give credit.

```{r osm}
osm2 <- opq(bbox = c(7.18783642,53.00393642,7.19693642,53.00953642)) %>%
    add_osm_features(features = list("natural"= "water", "highway" = "!primary")) %>%
  # add_osm_feature(key = "highway") %>%
  osmdata_sf()

ggplot()+
  geom_sf(data=osm2$osm_polygons)+
  geom_sf(data=osm2$osm_lines)+
  # scale_y_reverse()+
  theme_minimal()+
  theme(panel.grid = element_blank(),
        axis.text= element_blank(),
        panel.background = element_rect(
        fill = "#fcfcf4",
        colour = NA))+
  ggspatial::annotation_north_arrow(
    location = "tr", which_north = "true",
    pad_x = unit(0.75, "in"), pad_y = unit(3, "in"),
    style = ggspatial::north_arrow_nautical(
      line_col = "grey20",
      text_family = "ArcherPro Book"
    )
  )+
  labs(title="The Stellar Fort Bourtange",subtitle="Star-shaped Landscape of Bourtange, Netherlands",
       caption = "Source | OpenStreetMaps - Overpass API")



ggsave(filename="maps/15-osm.png",units="cm",width=24,height=18,dpi=150,device="png")
```

# 16	Oceania	Region made up of thousands of islands throughout the Central and South Pacific

# 17	Flow	Flow of transport, people


# 18	Atmosphere	Conditions of the atmosphere can be either weather or climate

# 19	5-minute map	Spend no more than 5 minutes to prepare this map

```{r 5_min}
# us_wtr <- st_join(waterway%>%st_transform(st_crs(us_area)),us_area,intersection=st_within)

tsunami<-inBoundsEQ[inBoundsEQ$tsunami==1,]
tsunami$dist2Town<- as.numeric(str_split(inBoundsEQ$place,"km",simplify=T)[1])

mD <- max(tsunami$dist2Town)

ggplot()+
  geom_sf(data=cont_us_area,fill=GRAY9,color="white")+
  geom_sf(data=tsunami[tsunami$tsunami==0,],color=GRAY7,aes(size=dist2Town/mD))+
  geom_sf(data=tsunami[tsunami$tsunami==1,],color="skyblue",aes(size=dist2Town/mD))+
  theme_test()+
  theme(axis.title.x = element_blank(),
        axis.text.x = element_blank(),
        axis.title.y = element_blank(),
        axis.text.y = element_blank(),
        # axis.ticks.x = element_blank(),
        # axis.ticks.y = element_blank(),
        panel.background = element_rect(
        fill = "#fcfcf4",
        colour = NA),
        # allows html in labs
        plot.subtitle = element_markdown(lineheight = 1.1,hjust = 0)
        )+
  labs(caption = "Source | USGS Earthquakes Catalog\nTimothy Oliver",
       title = "<span style='font-size:11pt'>2023
       <span style='color:#87CEEB;'>Tsunamis</span>",
       subtitle = "Better Version in Source Code")

ggsave(filename="maps/19-fiveMinutes.png",units="cm",width=24,height=18,dpi=150,device="png")
```

```{r better_version}
# us_wtr <- st_join(waterway%>%st_transform(st_crs(us_area)),us_area,intersection=st_within)

tsunami<-inBoundsEQ[inBoundsEQ$tsunami==1,]
dist2Town <- c()
for (place in tsunami$place){
  print(place)
  dist <- as.numeric(str_split(place,"km",simplify=T)[1])
  dist2Town <- c(dist2Town,dist)
}
tsunami$dist2Town<- dist2Town

mD <- max(tsunami$dist2Town)

ggplot()+
  geom_sf(data=cont_us_area,fill=GRAY9,color="white")+
  geom_sf(data=inBoundsEQ[inBoundsEQ$tsunami==0,],color=GRAY7,size=0.4)+
  geom_sf(data=tsunami[tsunami$tsunami==1,],color="skyblue",aes(size=dist2Town))+
  theme_test()+
  theme(axis.title.x = element_blank(),
        axis.text.x = element_blank(),
        axis.title.y = element_blank(),
        axis.text.y = element_blank(),
        # axis.ticks.x = element_blank(),
        # axis.ticks.y = element_blank(),
        panel.background = element_rect(
        fill = "#fcfcf4",
        colour = NA),
        # allows html in labs
        plot.title = element_markdown(lineheight = 1.5,hjust = 0)
        )+
  labs(caption = "Source | USGS Earthquakes Catalog\nTimothy Oliver",
       title = "<span style='font-size:11pt'>2023
       <span style='color:#87CEEB;'>Tsunamis</span>")+
  guides(fill=guide_legend(title="Miles to Town\nFrom Tsunami Epicenter"))

ggsave(filename="maps/19-fiveMinutes+debug.png",units="cm",width=24,height=18,dpi=150,device="png")
```

# 20	Outdoors	Map of mountains, trails, or something completely different

```{r outdoors}

ggplot()+
  geom_sf(data=us_area%>%filter(NAME=="Pennsylvania")%>%st_transform(3362),color="white",fill=GRAY9)+
  geom_sf(data=paParks[is.na(paParks$TYPE)==F,]%>%st_transform(3362),aes(color=TYPE),linewidth=1.4)+
  theme_test()+
  theme(axis.title.x = element_blank(),
        axis.text.x = element_blank(),
        axis.title.y = element_blank(),
        axis.text.y = element_blank(),
        # axis.ticks.x = element_blank(),
        # axis.ticks.y = element_blank(),
        panel.background = element_rect(
        fill = "#fcfcf4",
        colour = NA),
        )+
  labs(caption = "Source | PA Department of Conservation & Natural Resources",
       title = "Pennsylvania Parks and Preserves")+
  guides(fill=guide_legend(title="Park Type"))

ggsave(filename="maps/20-outdoors.png",units="cm",width=24,height=18,dpi=150,device="png")
```

# 21	Raster	Pixels, please
```{r dead_sea}
png(filename="maps/21-raster.png",height=960,width=1024,units="px")
# coordinates(dead_sea) <- ~x+y
# bubble(dead_sea, "elev", col=c("#a8f490","#06ffdb"),main="elevation (m)")
# 
# gridded(dead_sea) = T
# image(dead_sea["elev"])
# title("Distance to River (red=0)")
# 
# elev.idw = idw(elev~1,meuse,meuse.grid)
# 
# spplot(elev.idw["var1.pred"],main = "Elevation Inverse Distance Weighted Interpolations")
library(gstat)
data(meuse)
coordinates(meuse) = ~x+y
# bubble(meuse, "elev", col=c("#a8f490","#06ffdb"),main="elevation (m)")
data(meuse.grid)
coordinates(meuse.grid) = ~x+y
gridded(meuse.grid) = T
# image(meuse.grid["dist"])
elev.idw = gstat::idw(elev~1,meuse,meuse.grid)
lel.vgm=variogram(log(elev)~1,meuse)
lel.fit = fit.variogram(lel.vgm,model=vgm(1,"Sph",900,1))
lel.kriged = krige(log(elev)~1, meuse, meuse.grid, model = lel.fit)

spplot(lel.kriged["var1.pred"])
library(grid)
grid.text("The High Points of the Meuse River Valley (m)", y=unit(0.96, "npc"))

dev.off()


```

# 22	North is not always up	⬆️
```{r north_is_left}
q <- opq(bbox = "parque o'higgins, santiago") %>%
    add_osm_feature(key = 'highway') %>% 
  osmdata_sf()

ggplot()+
  geom_sf(data=q$osm_polygons)+
  geom_sf(data=q$osm_lines)+
  # scale_y_reverse()+
  theme_minimal()+
  theme(panel.grid = element_blank(),
        axis.text= element_blank())+
  ggspatial::annotation_north_arrow(
    location = "tr", which_north = "true",
    pad_x = unit(0.25, "in"), pad_y = unit(0, "in"),
    style = ggspatial::north_arrow_nautical(
      fill = c("grey40", "white"),
      line_col = "grey20",
      text_family = "ArcherPro Book"
    )
  )+
  labs(title="Anywhere but North for Attractions",subtitle="Parque o'Higgins - Santiago, Chile",
       caption = "Source | USGS Earthquakes Catalog")

# ggsave(filename="maps/22-notUp.png",units="cm",width=24,height=18,dpi=150,device="png")
```

# 23	3D	The third dimension

```{r some_shade}
png(filename="maps/23-3d.png",height=960,width=1024,units="px")
mf_shadow(cont_us_area)
mf_map(cont_us_area, add = TRUE)
# plot baseball values
mf_map(
  x = st_as_sf(inBoundsEQ),
  var = "mag",
  type = "grad",
  inches = 0.00034,
  alpha = 0.3,
  leg_pos = "bottomright2",
  leg_title = "Earthquake Magnitude"
)
# layout
mf_layout(
  title = "2023 Earthquakes",
  credits = paste0(
    "Sources: USGS Earthquake Catalog\n",
    "mapsf ",
    packageVersion("mapsf")
  )
)

dev.off()

# ggsave(filename="maps/23-3d.png",units="cm",width=24,height=18,dpi=150,device="png")
```

# 24	Black & white	Only two colors allowed

```{r black_and_white}
osm3 <- opq(bbox = "central park") %>%
    # add_osm_feature(key = "natural") %>% 
  add_osm_feature(key = "highway") %>%
  osmdata_sf() 

ggplot()+
  geom_sf(data=osm3$osm_polygons%>% st_transform(crs="EPSG:2260"))+
  geom_sf(data=osm3$osm_lines%>% st_transform(crs="EPSG:2260"))+
  # scale_y_reverse()+
  theme_minimal()+
  theme(panel.grid = element_blank(),
        axis.text= element_blank())+
  ggspatial::annotation_north_arrow(
    location = "tr", which_north = "true",
    pad_x = unit(0.75, "in"), pad_y = unit(3, "in"),
    style = ggspatial::north_arrow_nautical(
      line_col = "grey20",
      text_family = "ArcherPro Book"
    )
  )+
  labs(title="The Island in the Grid",subtitle="Central Park, NY",
       caption = "Source | OpenStreetMaps - Overpass API")



ggsave(filename="maps/24-black_and_white.png",units="cm",width=24,height=18,dpi=150,device="png")
```

# 25	Antarctica	A cold continent that can't be reached with Web Mercator

# 26	Minimal	Less is more
```{r minimal_road_density}



ggplot()+
  geom_sf(data = cont_us_area, fill = "lightblue",col = "white")+
  geom_sf(data = states[states$n == 1,],
          fill="pink",color="white")+
  geom_sf(data = states[states$n > 1,],
          fill="#FF4738",color="white")+
  # geom_sf(data=inBoundsEQ,color=GRAY9,size=0.2)+
  labs(caption = "Note: earthquakes of magnitude 2.5 or greater shown\n\nSource | USGS Earthquakes Catalog\nTimothy Oliver",
       title= "<span style='font-size:11pt'>Earthquake
       <span style='color:#ADD8E6;'>Free</span> to Earthquake
       <span style='color:#FF4738;'>Filled</span> States")+
  theme_test()+
  theme(axis.title.x = element_blank(),
        axis.text.x = element_blank(),
        axis.title.y = element_blank(),
        axis.text.y = element_blank(),
        # axis.ticks.x = element_blank(),
        # axis.ticks.y = element_blank(),
        panel.background = element_rect(
        fill = "#fcfcf4",
        colour = NA),
        # allows html in labs
        plot.title = element_markdown(lineheight = 2.5,hjust = 0)
        )




ggsave(filename="maps/26-minimal.png",units="cm",width=24,height=18,dpi=150,device="png")


```

# 27	Dot	Dot density, a single dot in space or something different

# 28	Is this a chart or a map?	In thematic maps, you can't always tell. Try to stretch the limits

# 29	Population	A classic theme for a map

# 30	"My favorite.."	You choose!



```{r tsunami_eq_CA,cache=T}
CA <- us_area%>%filter(NAME=="California")
buf <-st_buffer(st_centroid(CA),dist=15600)
# cfEQ <- st_join(earthquakes_23,
#         buf,
#         join=st_within) 
# 
# nrow(cfEQ)
# nrow(earthquakes_23)

```

```{r map_simple_data}
ggplot()+
  geom_sf(data=ocean,fill=GRAY9,color="white")+
  geom_sf(data=birds,color='pink')

ggplot()+
  geom_sf(data=CA%>%st_transform("EPSG:3665"))+
  geom_sf(data=buf,color='black',fill='lightblue')

ggplot()+
  geom_sf(data=q$osm_polygons)+
  geom_sf(data=q$osm_lines)+
  # scale_y_reverse()+
  theme_minimal()+
  theme(panel.grid = element_blank(),
        axis.text= element_blank())+
  ggspatial::annotation_north_arrow(
    location = "tr", which_north = "true",
    pad_x = unit(0.25, "in"), pad_y = unit(0, "in"),
    style = ggspatial::north_arrow_nautical(
      fill = c("grey40", "white"),
      line_col = "grey20",
      text_family = "ArcherPro Book"
    )
  )
  
```

```{r knit_etc., eval=F, include=F}
rmarkdown::render(input="all_maps.Rmd",output_format = "html_document",output_file = "The Thirty-Day Mapping Journey.html")
```